---
title: "Assignment 4"
author: "Ishtiaq Mahmud Fahim"
format: pdf
editor: visual
---

# 6.1 Solution

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(alr4)
```

\begin{center}

NH : lifeExp $\sim$ 1

AH : lifeExp $\sim$ group

\end{center}

```{r}
dat <- UN11 %>% 
  select(lifeExpF, group)

dat <- dat %>% 
  mutate(
    U2 = if_else(group == "other", 1, 0),
    U3 = if_else(group == "africa", 1, 0),
  )

n <- length(dat$group)

Y <- matrix(
  c(dat$lifeExpF),
  ncol = 1
)
X <- matrix(
  c(
    rep(1,n),
    dat$U2,
    dat$U3
  ),
  ncol = 3
)

H <- X %*% solve(t(X) %*% X) %*% t(X)
I <- diag(1, 199, 199)
J <- matrix(1, 199, 199)


RSS_0 <- t(Y) %*%  (I - J/199) %*% Y
df_0 <- 198

RSS_1 <- t(Y) %*%  (I - H) %*% Y
df_1 <- 196

F_cal <- ((RSS_0 - RSS_1)/(df_0 - df_1))/(RSS_1 / df_1)
F_cal

p_val <- pf(F_cal, 2, df_1, lower.tail = FALSE)
cp <- if_else(p_val < 0.001, "<0.001",as.factor(p_val))
cp
```

-   The F-test Yielded an F-value of **159.2687** with a p-value less than **0.001**

-   So it suggests that the factor in the model significantly improves the fit compared to a model without it. Therefore, we may reject the null hypothesis in favor of the alternative hypothesis.

\newpage

# 6.4 Solution

\begin{center}

NH : lifeExp $\sim$ log(ppgdp) + group $\colon$ log(ppgdp)

AH : lifeExp $\sim$ group + log(ppgdp) + group $\colon$ log(ppgdp)

\end{center}

## 6.4.1

The **NH** model states that female life expectancy depends on the **log(ppgdp)** and the **effect of ppgdp varies across groups**. Also the **NH** model violets the **marginality principle** because it includes an interaction term **group : log(ppgdp)** without including the corresponding main effect for group.

## 6.4.2

```{r}

dat2 <- UN11 %>% 
  mutate(
    log_ppgdp = log(ppgdp)
  ) %>% 
  select(lifeExpF, group, log_ppgdp)

dat2 <- dat2 %>% 
  mutate(
    U2 = if_else(group == "other", 1, 0),
    U3 = if_else(group == "africa", 1, 0),
    U2x = U2*log_ppgdp,
    U3x = U3*log_ppgdp
  )

n <- length(dat2$log_ppgdp)

Y <- matrix(
  c(dat2$lifeExpF),
  ncol = 1
)

X0 <- matrix(
  c(
    rep(1,n),
    dat2$log_ppgdp,
    dat2$U2x,
    dat2$U3x
  ),
  ncol = 4
)

X1 <- matrix(
  c(
    rep(1,n),
    dat2$U2,
    dat2$U3,
    dat2$log_ppgdp,
    dat2$U2x,
    dat2$U3x
  ),
  ncol = 6
)

H0 <- X0 %*% solve(t(X0) %*% X0) %*% t(X0)

H1 <- X1 %*% solve(t(X1) %*% X1) %*% t(X1)


I <- diag(1, 199, 199)



RSS_0 <- t(Y) %*%  (I - H0) %*% Y
df_0 <- 195

RSS_1 <- t(Y) %*%  (I - H1) %*% Y
df_1 <- 193

F_cal <- ((RSS_0 - RSS_1)/(df_0 - df_1))/(RSS_1 / df_1)
F_cal

p_val <- round(pf(F_cal, 5, df_1, lower.tail = FALSE),5)
cp <- if_else(p_val < 0.001, "<0.001",as.factor(p_val))
cp
```

## 

The F-test statistic of **2.93** with a p-value of **0.014** indicates that there is a statistically significant difference between the **NH** and **AH** models at the 5% significance level. This suggests that adding the main effect of **group** improves the model.
