---
title:  "Assignment1"
author: "Ishtiaq Mahmud Fahim"
format: pdf
editor: visual
---

# 2.2 Solution

```{r pull, warning=F, message=F}
library(alr4)
library(tidyverse)

mydat <- as_tibble(UBSprices %>% 
  mutate(
    city = row.names(UBSprices)
  ) %>% 
  select(city, rice2003, rice2009)
)
ggplot(mydat)+
  geom_point(
    aes(
      x = rice2003,
      y = rice2009
    )
  )+
  geom_abline(intercept = 0, slope = 1, color = "blue", linetype = "solid",
               size = 1
  )+
  geom_smooth(
    aes(
      x = rice2003,
      y = rice2009
    ),
    se = F,
    col = "red",
    method = "lm"
  )
```

-   The **Blue Line** represents the **y = x** line.

-   The **Red Line** represents the **Mean** line.

### 2.2.1

The line **y = x** represents a situation where the rice price in 2009 is equal to the rice price in 2003 for each city.

-   Points **above** the line indicate cities where rice prices were **higher in 2009** compared to 2003

-   Points **below** the line show cities where prices were **lower in 2009** compared to 2003

\newpage

### 2.2.2

```{r}

mydat %>% 
  mutate(
    price_difference = rice2009 - rice2003
  ) %>% 
  filter(price_difference == max(price_difference) |
         price_difference == min(price_difference))

```

-   **Vilnius** has the largest price increase.

-   **Mumbai** has the largest price decrease.

### 2.2.3

```{r}
sumsq <- function(x,y){
  ss <- sum((x-mean(x))*y)
  return(ss)
}

slop <- function(x,y){
  sumsq(x,y)/sumsq(x,x)
}

slop(mydat$rice2003,mydat$rice2009)
```

The OLS line where $\hat{y} = \hat{\beta_0} + \hat{\beta_1}x$ , $\hat{\beta_1} = 0.5013831$ suggests that for each unit **increase in the 2003 price**, the **2009 price increases by less than one, around 0.501 unit on average**.

This implies a **lower rate of increase** in 2009 prices relative to 2003, which could suggest that overall prices may not have increased as much from 2003 to 2009 but **does not** suggest that the prices are lower in 2009 than in 2003.

\newpage

```{r, ref.label= "pull", results=T, echo=FALSE, message=FALSE, warning=FALSE}

```

### 2.2.4

1.  **Non-Linearity :** The relationship between 2003 and 2009 prices may not be linear, as economic factors influencing prices might vary across regions.

2.  **Outliers :** Some cities might have unique economic conditions causing them to deviate significantly from the trend, impacting the regression line's accuracy.

\newpage

# 2.9 Solution

### 2.9.1

$$
E(Y \mid X=x) = \beta_0 + \beta_1x
$$

The value of the predictor **X** is replaced by $Z = aX + b ; a\not=0$

This transformation leads to a new regression model in terms of **Z** :

$$
Y = \gamma_0 + \gamma_1z + \epsilon
$$

By expressing **X** in terms of **Z** it can be achieved that;

$$
X = \frac{Z-b}{a}
$$

$$
E(Y \mid Z=z) = (\beta_0 - \frac{\beta_1b}{a}) + \frac{\beta_1}{a}z
$$

Hence,

$$
\gamma_0 = \beta_0 - \frac{\beta_1b}{a}
$$

and

$$
\gamma_1 = \frac{\beta_1}{a}
$$

Since the Response **Y** has not change, the estimate of $\sigma^2$ remains **unchanged**.

Because the transformation of **X** does not change the inherent spread or variability of **Y** around the regression line. The variance of **Y** given **X** is a property of **Y**'s distribution in relation to **X** and is not altered by scale of location of **X** itself.

Now for **t-statistic** :

$$
t_{\gamma_1} = \frac{\hat{\gamma_1}}{\hat{SE(\hat{\gamma_1})}} = \frac{\frac{\hat{\beta_1}}{a}}{\frac{\hat{SE(\hat{\beta_1})}}{|a|}} = \frac{\hat{\beta_1}}{SE(\hat{\beta_1})}
$$

So, it remains the same.

\newpage

### 2.9.2

$$ E(Y \mid X=x) = \beta_0 + \beta_1x $$

The value of the response **Y** is replaced by $V = dY ; d \not= 0$

This transformation leads to a new regression model in terms of **V** :

$$
V = \delta_0 + \delta_1x + \epsilon
$$

By expressing **V** in terms of **Y** it can be achieved that;

$$
E(dY \mid X=x) = d\beta_0 + d\beta_1x
$$

Hence,

$$
\delta_0 = d\beta_0
$$

and

$$
\delta_1 = d\beta_1
$$

The variance changes as the response **Y** has changed,

$$
Var[V\mid X=x] = Var[dY\mid X=x] = d^2Var[Y\mid X=x]
$$

So,

$$
\hat{\sigma^2_v} = d^2\hat{\sigma^2}
$$

Now for **t-statistic :**

$$
t_{\delta_1} = \frac{\hat{\delta_1}}{\hat{SE(\hat{\delta_1})}} = \frac{d\hat{\beta_1}}{\hat{SE(d\hat{\beta_1})}} = \frac{d\hat{\beta_1}}{d\hat{SE(\hat{\beta_1})}} = \frac{\hat{\beta_1}}{\hat{SE(\hat{\beta_1})}} = t_{\beta_1}
$$

So, it remains the same.
